Vitis amurensis LM

### 2018-10-29 V_amurensis_LM.Rmd
V. amurensis prep

```{r setup}
library(readxl)
library(dplyr)
library(skimr)
library(ggplot2)
library(gganimate)
library(segmented)
Vitis_90_ratios <- read_excel("R_DiLP_all_species.xlsx", 
     sheet = "all.90", col_types = c("text", 
         "numeric", "numeric", "numeric", 
         "skip", "skip", "skip", "skip", 
         "skip", "skip", "skip", "skip", 
         "skip", "skip", "skip", "numeric", 
         "skip", "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric"))
```

@@

We first tested each of the variables for *Vitis acerifolia*. Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).

We next tested each of the variables for *Vitis aestivalis*. Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).

We next tested each of the variables for *Vitis amurensis*. Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).

We finally tested each of the variables for *Vitis riparia*. Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).

@@

We need the means for each leaf character for both years combined.
```{r}
amurensis <-filter(Vitis_90_ratios, species == "amurensis")
am_l_means <-  amurensis %>% 
  group_by(leaf) %>% 
  summarize(avg.tt = mean(total.teeth),
            avg.fdr = mean(feret.diameter.ratio),
            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
            avg.ta = mean(average.tooth.area, na.rm = TRUE),
            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
            avg.tp = mean(teeth.perimeter),
            avg.tip = mean(teeth.int.perimeter),
            avg.pa = mean(perimeter.area),
            avg.pr = mean(perimeter.ratio),
            avg.cp = mean(compactness),
            avg.sf = mean(shape.factor),
            avg.tba = mean(teeth.blade.area))
```

Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF), so we need to split these variables for each year.
```{r}
am_13 <- filter(amurensis, year == 2013)
am_15 <- filter(amurensis, year == 2015)
```

@@

Total teeth:perimeter
```{r}
am_tp13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.tp = mean(teeth.perimeter, na.rm = TRUE))

am_tp15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.tp = mean(teeth.perimeter, na.rm = TRUE))
```

Perimeter:area
```{r}
am_pa13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.pa = mean(perimeter.area, na.rm = TRUE))

am_pa15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.pa = mean(perimeter.area, na.rm = TRUE))
```

>>>>

```{r setup}
library(readxl)
library(dplyr)
library(skimr)
library(ggplot2)
library(gganimate)
library(segmented)
Vitis_90_ratios <- read_excel("R_DiLP_all_species.xlsx", 
     sheet = "all.90", col_types = c("text", 
         "numeric", "numeric", "numeric", 
         "skip", "skip", "skip", "skip", 
         "skip", "skip", "skip", "skip", 
         "skip", "skip", "skip", "numeric", 
         "skip", "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric"))
amurensis <-filter(Vitis_90_ratios, species == "amurensis")
```

@@

We next tested each of the variables for *Vitis amurensis*.

```{r}
t.test(total.teeth ~ year, data= amurensis, var.equal= TRUE)
t.test(feret.diameter.ratio ~ year, data= amurensis, var.equal= TRUE)
t.test(tooth.area.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(tooth.area.int.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(average.tooth.area ~ year, data= amurensis, var.equal= TRUE)
t.test(tooth.area.blade.area ~ year, data= amurensis, var.equal= TRUE)
t.test(teeth.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(teeth.int.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(perimeter.area ~ year, data= amurensis, var.equal= TRUE)
t.test(perimeter.ratio ~ year, data= amurensis, var.equal= TRUE)
t.test(compactness ~ year, data= amurensis, var.equal= TRUE)
t.test(shape.factor ~ year, data= amurensis, var.equal= TRUE)
t.test(teeth.blade.area ~ year, data= amurensis, var.equal= TRUE)
```
Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), perimeter to ratio (PR), and shape factor (SF).

@@

We need the means for each leaf character for both years combined.
```{r}
am_l_means <-  amurensis %>% 
  group_by(leaf) %>% 
  summarize(avg.tt = mean(total.teeth, na.rm = TRUE),
            avg.fdr = mean(feret.diameter.ratio, na.rm = TRUE),
            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
            avg.ta = mean(average.tooth.area, na.rm = TRUE),
            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
            avg.tp = mean(teeth.perimeter, na.rm = TRUE),
            avg.tip = mean(teeth.int.perimeter, na.rm = TRUE),
            avg.pa = mean(perimeter.area, na.rm = TRUE),
            avg.pr = mean(perimeter.ratio, na.rm = TRUE),
            avg.cp = mean(compactness, na.rm = TRUE),
            avg.sf = mean(shape.factor, na.rm = TRUE),
            avg.tba = mean(teeth.blade.area, na.rm = TRUE))
```

Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), perimeter to ratio (PR), and shape factor (SF), so we need to split these variables for each year.
```{r}
am_13 <- filter(amurensis, year == 2013)
am_15 <- filter(amurensis, year == 2015)
```

@@

Perimeter ratio
```{r}
am_pr13_mean <- amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))

am_pr15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))
```

Shape factor
```{r}
am_sf13_mean <- amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))

am_sf15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
```

We tested leaf number and feret diameter ratio.
```{r}
am_fdr_plot <- ggplot(amurensis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
am_fdr_plot <- am_fdr_plot +labs(x = "Leaf number",
                               y = "FDR")
am_fdr_plot

```

```{r}
am_fdr_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
am_fdr_mean_plot <- am_fdr_mean_plot +labs(x = "Leaf number",
                               y = "FDR")
am_fdr_mean_plot
```

We ran a LM for leaf number and feret diameter ratio.
```{r}
am_fdr_lm <- lm(avg.fdr ~ leaf, data = am_l_means)
summary(am_fdr_lm)
```

```{r}
am_fdr_seg <- segmented(am_fdr_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(4, 7, 11)))
summary(am_fdr_seg)
```

```{r}
am_fdr_seg$psi
slope(am_fdr_seg)
am_fdr_fitted <- fitted(am_fdr_seg)
am_fdr_model <- data.frame(Leaf = am_l_means$leaf, FDR = am_fdr_fitted)
ggplot(am_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
am_fdr_lines <- am_fdr_seg$psi[, 2]
am_fdr_mean_break <- am_fdr_plot + geom_vline(xintercept = am_fdr_lines, linetype = "dashed")
am_fdr_mean_break
ggsave("am_fdr_break.png")
```
